Work Fragmentation in Developer Interaction Data

### 2015-12-17 paper.tex
Reviews 1.3, 1.14, 2.1, 3.8

@@

@@

We did so by computing Cohen's $d$ on the quantity of edits before and after the interruption. To obtain a significant effect, we need the presence of activity both before and after. Not all the interruptions meet this criteria however: some are located close to the start or the end of a session, or too close to another interruption.  In total, 53\% of the interruptions had 30 minutes before and after the interruption and were selected for the analysis. %We used the Algorithm \ref{alg:local_effect} for the classification, and the 
Table \ref{tbl:local_effect} shows the applied thresholds and the results, accompanied with a typical example of an interruption in each category. This local analysis shows that there are indeed three well-defined groups of interruptions, with the two largest of them having clear effects on the activity in the session.

@@

\begin{table}[ht!]
	\small
	\renewcommand{\arraystretch}{1.3}
	\caption{Statistics about the sessions in UDC. }
	\label{tbl:stats_sessions_udc}
	\centering
	\begin{tabular}{| p{3cm} |p{3cm} |} 
		\hline 
		stat & value \\  
		\hline 
		\hline 
		total & 15,825 \\
		average duration & 256 min \\
		max duration & 3,607 min \\
		min duration & 30 min \\
		std. deviation & 229 min \\
		\hline

@@

\begin{figure}[t]
	\centering
	\begin{tabular}{c}
		\includegraphics[width=0.6\linewidth,clip=, angle=0]{figures/hist_session_duration_udc} 
	\end{tabular}
	\caption{Histogram of the duration of the sessions in UDC.}
	\label{fig:hist_session_duration_udc}
\end{figure}

@@

\begin{table}[ht!]
\caption{Local Effect of Interruption. }
\label{tbl:local_effect_udc}
\centering
\begin{tabular}{p{4cm} | c}
effect & pattern \\
\hline
\textit{negative} (44\%): when the frequency of edit events decreases after the interruption
	& \includegraphics[valign=m,scale=0.25]{figures/udc_neg_ints} \\
\textit{positive} (44\%): when the frequency of edit events increases after the interruption 
	& \includegraphics[valign=m,scale=0.25]{figures/udc_pos_ints} \\
\textit{neutral} (12\%): when there is no well defined effect before or after the interruption 
	& \includegraphics[valign=m,scale=0.25]{figures/udc_neu_ints} 
\end{tabular}
\end{table}

@@

\subsection{Recovery time after positive and negative interruptions}
We take a closer look into the recovery time by splitting it into interruptions of positive and negative events. As expected the edit ratio is higher during recovery time after a positive interruption and the contrary is seen after a negative interruption. The Figure \ref{fig:box_recovery_effect_udc} show a comparison of both effects and the overall edit ratio considering all interruptions.

@@

We can see that after a positive interruption the edit ratio during recovery time is much higher than after a negative interruption, which agrees with the results of the local analysis of interruptions and the observed effects on the metrics. %The reason of this change is probably due to the difference on the activities performed after the interruption according to the effect

>>>>

After the transformation phase, we had the sessions in a proper format to take on the analysis. Some statistics about the sessions in Mylyn can be seen in the Table \ref{tbl:stats_sessions_mylyn}, and the distribution of the productive time of this sessions is shown in the Figure \ref{fig:mylyn_trace_duration}. Both the Table and Figure consider the productive time of the session, that is the total time without considering the time consumed by interruptions. Also, the outliers are ignored by setting a threshold of three times the standard deviation. This filter is used only to show the distribution, to avoid a very long tail provoked by unusually long sessions.

\begin{table}[ht!]
	\tiny
	\renewcommand{\arraystretch}{1.3}
	\caption{Statistics about the sessions in Mylyn. }
	\label{tbl:stats_sessions_mylyn}
	\centering
	\begin{tabular}{| p{3cm} |p{3cm} |} 
		\hline 
		stat & value \\  
		\hline 
		\hline 
		total & 4,284 \\
		min. & 2 min \\
		1st Qu. & 18 min\\
		median & 35 min \\
		mean & 48.91 min \\
		3rd Qu. & 62 min \\
		max. & 535 min\\
		\hline

\begin{figure}[t]
	\centering
	\begin{tabular}{c}
		\includegraphics[width=0.6\linewidth,clip=, angle=0]{figures/udc_trace_duration} 
	\end{tabular}
	\caption{Histogram of the duration of the sessions in Mylyn.}
	\label{fig:mylyn_trace_duration}
\end{figure}

@@

We measure the effects on productivity using the edits and selections per minute as indicators. The edit ratio might be also a good indicator, under the assumption that a productive programmer will perform more coding activities than selections or navigations. However, this is not always the case.

@@

We did so by computing Cohen's $d$ on the quantity of edits before and after the interruption. To obtain a significant effect, we need the presence of activity both before and after. Not all the interruptions meet this criteria however: some are located close to the start or the end of a session, or too close to another interruption.  In total, 53\% of the interruptions had 30 minutes before and after the interruption and were selected for the analysis. The rest 47\% of the interruptions can not be used on this analysis due to the lack of time around them. %We used the Algorithm \ref{alg:local_effect} for the classification, and the 
Table \ref{tbl:local_effect} shows the applied thresholds and the results, accompanied with a typical example of an interruption in each category. This local analysis shows that there are indeed three well-defined groups of interruptions, with the two largest of them having clear effects on the activity in the session.

@@

\begin{table}[ht!]
	\tiny
	\renewcommand{\arraystretch}{1.3}
	\caption{Statistics about the sessions in UDC. }
	\label{tbl:stats_sessions_udc}
	\centering
	\begin{tabular}{| p{2cm} |p{2cm} |} 
		\hline 
		stat & value \\  
		\hline 
		\hline

		total & 15,825 \\
		min. & 2 min \\
		1st Qu. & 11 min \\
		median & 22 min \\
		mean & 31.09 min \\
		3rd Qu. & 45 min \\
		max. & 131 min \\
		\hline

@@

\begin{figure}[t]
	\centering
	\begin{tabular}{c}
		\includegraphics[width=0.6\linewidth,clip=, angle=0]{figures/udc_trace_duration} 
	\end{tabular}
	\caption{Histogram of the duration of the sessions in UDC.}
	\label{fig:hist_session_duration_udc}
\end{figure}

@@

\begin{table}[ht!]
\caption{Local Effect of Interruption. }
\label{tbl:local_effect_udc}
\centering
\begin{tabular}{p{4cm} | c}
effect & pattern \\
\hline
\textit{negative} (44\%): when the frequency of edit events decreases after the interruption
	& \includegraphics[valign=m,scale=0.25]{figures/udc_neg_ints_2} \\
\textit{positive} (44\%): when the frequency of edit events increases after the interruption 
	& \includegraphics[valign=m,scale=0.25]{figures/udc_pos_ints_2} \\
\textit{neutral} (12\%): when there is no well defined effect before or after the interruption 
	& \includegraphics[valign=m,scale=0.25]{figures/udc_neu_ints_2} 
\end{tabular}
\end{table}

@@

We saw in RQ3 different patterns of activity around interruptions when they had positive or negative effects; the edits per minute are higher after a positive interruption and the contrary after a negative one. The effect is more marked in the immediate minutes after the interruption, as can be seen in the Figure \ref{fig:box_recovery_effect_udc}.

@@
